P.berghei pir analysis

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Doing the berghei df melting and averging b4 top_n fun

@@

```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
top_n_genes <- function(tpm_input, 
                        n = 5, 
                        threshold = 0){
  tpm = tpm_input$tpm %>% 
    #Melt the data frame for using dplyr
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1))) %>% 
    group_by(Geneid, stage) %>%  
    dplyr::summarise(median_tpm = median(tpm),
                     mean_tpm = mean(tpm), 
                     min_tpm = min(tpm), 
                     max_tpm = max(tpm), 
                     sd_tpm = sd(tpm)) %>% 
    group_by(stage) %>% 
    mutate(experiment = tpm_input$experiment)

  #Calculating the index used to make the kable table

  kable_index = rep(n, times = length(unique(tpm$stage)))
  names(kable_index) = unique(tpm$stage)

  all_genes = tpm %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue")))
  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
                                                                allgene.info$Geneid)]
  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
                                                         allgene.info$Geneid)]
  all_genes = all_genes %>% 
    kable(format = 'html', 
          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed total genes'), 
          escape = FALSE) %>%
    kable_styling() %>% 
    pack_rows(index = kable_index)

  pir_genes = tpm %>% 
    filter(Geneid %in% pir_id) %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
    mutate(Geneid = as.character(Geneid)) %>% 
    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
                                                               allgene.info$Geneid)]) %>% 
    mutate(Geneid = ifelse(grepl(Geneid,
                                 pattern = 'PBANKA_0100500'),
                           cell_spec(Geneid, bold = TRUE),
                           Geneid)) %>%
    kable(format = 'html', 
          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed pir genes'), 
          escape = FALSE) %>% 
    pack_rows(index = kable_index) %>% 
    kable_styling()

  # print(paste0('Completed analysis of ', tpm_input$experiment))
  list = list(tpm, all_genes, pir_genes, tpm_input$experiment)
  names(list) = c('tpm', 'all_genes', 'pir_genes', 'experiment')
  return(list)
}

@@

all_top <- lapply(expt_tpms, 
                  top_n_genes)

@@

big_top = bind_rows(lapply(all_top,function(x){
  data = x$tpm
  data$stage = paste(x$experiment, data$stage, sep = '_')
  data
}))

>>>>

```{r melt_df}
berghei_tpm_melt <- berghei_tpm %>% 
    #Melt the data frame for using dplyr
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    mutate(experiment = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1)),
           stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,2)))

berghei_tpm_avg_byexpt <- berghei_tpm_melt %>% 
  group_by(Geneid, experiment, stage) %>%  
  dplyr::summarise(median_tpm = median(tpm),
                   mean_tpm = mean(tpm), 
                   min_tpm = min(tpm), 
                   max_tpm = max(tpm), 
                   sd_tpm = sd(tpm)) 
```

@@

```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
top_n_genes <- function(tpm_avg, 
                        n = 5, 
                        threshold = 0){
  #Calculating the index used to make the kable table

  kable_index = rep(n, times = length(unique(tpm_avg$stage)))
  names(kable_index) = unique(tpm_avg$stage)

  all_genes = tpm_avg %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue")))
  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
                                                                allgene.info$Geneid)]
  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
                                                         allgene.info$Geneid)]
  all_genes = all_genes %>% 
    kable(format = 'html', 
          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed total genes'), 
          escape = FALSE) %>%
    kable_styling() %>% 
    pack_rows(index = kable_index)

  pir_genes = tpm_avg %>% 
    filter(Geneid %in% pir_id) %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
    mutate(Geneid = as.character(Geneid)) %>% 
    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
                                                               allgene.info$Geneid)]) %>% 
    mutate(Geneid = ifelse(grepl(Geneid,
                                 pattern = 'PBANKA_0100500'),
                           cell_spec(Geneid, bold = TRUE),
                           Geneid)) %>%
    kable(format = 'html', 
          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed pir genes'), 
          escape = FALSE) %>% 
    pack_rows(index = kable_index) %>% 
    kable_styling()

  # print(paste0('Completed analysis of ', tpm_input$experiment))
  list = list(all_genes, pir_genes, tpm_input$experiment)
  names(list) = c('all_genes', 'pir_genes', 'experiment')
  return(list)
}

@@

all_top <- lapply(unique(berghei_tpm_avg_byexpt$experiment), 
                  function(expt) top_n_genes(berghei_tpm_avg_byexpt[berghei_tpm_avg_byexpt$experiment == expt,]))

@@

big_top = bind_rows(lapply(all_top,function(x){
  data = x$tpm
  data$stage = paste(x$experiment, data$stage, sep = '_')
  data
}))
big_top <- berghei_tpm_melt
